Revisiting `Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation

### 2023-05-19 analysis/quarto/paper.tex
work on transpercy section

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public good game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, boxrule=0pt, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

Today's actions are tomorrow's result. There are many settings in which
current decisions affect future outcomes and with it, future decision
spaces. Opting for environmental friendly policies today not only
reduces carbon dioxide omissions immediately but also helps us to reach
the Paris climate targets tomorrow. Deferring these policies, may not
necessarily prevent us from reaching these targets, but it makes it more
difficult in the future. Hence, today's actions (or the omission
thereof) not only affect intermediate outcomes but also the number of
paths one can choose to reach certain goals. This apllies especially to
carbon dioxide emissions where its current stock will last for well over
a millennium \citep{Inman2008, CalzolariEtAl2018}.

@@

In fact, the experiment reported in this article qualifies for the label
\emph{`pre-registered'} at a first glance: Parts of the analyses were
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. In addition, I pre-registered the exact
analyses I planned to run when I designed the experiment on
\href{https://github.com/Howquez/coopUncertainty/blob/July21Replication/analysis/reports/rmd}{GitHub}.\footnote{https://github.com/Howquez/coopUncertainty/tree/July21Replication/analysis/reports/rmd
  to run the code, you need to executed the .Rmd files in this
  repository in the order that is indicated by its file names.}.

The analysis that made the most sense after having collected the data is
a little different though: Because the more representative subject pool
was exhausted earlier than expected, I recruited students which opened
up a new research direction: assessing generalizability. Hence, this
article combines both confirmatory as well as exploratory research.

Accordingly, I changed many of the analysis scripts.\footnote{The source
  code of this document contains the analysis and can be found
  \href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}{here}:
  https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}
However, because the originally planned analysis code is archived and
reproducible, these changes are transparent and easy to assess,
showcasing (a), (b) and the transition from (a) to (b).

@@

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific} reproduction\footnote{\citet{Parsons2022}
  use the term of a \emph{conceptual replication} which means the same.}
of one treatment of GMTV's dynamic public good game. The pure
reproduction re-analyzes the original data.
\protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{errors I identified} in the original paper. The scientific
replication, where I utilize a different sample drawn from a different
population in a different situation, is described in the following
sections.

@@

\hypertarget{sec-replication}{%
\subsection{Pre-registered GMTV Replication}\label{sec-replication}}

@@

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is the replication sample (consisting
of both students and non-students) more pro-social than the original
sample? A two-sided rank sum test reveals that it is not (p=0.3926).
Both samples contributed 10 tokens, that is, 50\% of their endowments on
average (median and mean). Moreover, both samples' initial contributions
resemble initial contributions participants usually make in the static
game with partner matching.\footnote{See Figure 3B in
  \citet[p.989]{fehrgaechter2000}, for instance.} However, in the
dynamic game presented here, we are particularly interested in the
subsequent periods because differences add up exponentially. Do the two
groups remain similar over the course of time?

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{no}.
Figure~\ref{fig-share-of-contributions} illustrates that the samples
make similar contributions at the beginning and the end of the game but
behave differently in between. More precisely, the left panel--depicting
the average contributions in absolute terms--shows that the original
sample contributed more than the replication sample \emph{in all but the
first and last period}.\footnote{In period five, this difference is
  significant: a two-sided rank sum test yields p=0.0486.} For this
reason, the original sample's behavior differs from the replication
sample's behavior in two aspects: it contributes more and exhibits a
considerable drop in the last period (whereas the replication sample's
contributions flatten).

Note that increasing contributions over time imply increasing endowments
over time. Hence, absolute contributions do not tell much about the
willingness to cooperate. For this reason, the right panel in
Figure~\ref{fig-share-of-contributions} shows the average \emph{share of
endowments contributed} over time. Both samples exhibit a similar
pattern: their share of endowments contributed declined and did not
stabilize. However, both samples also differ with respect to one aspect:
the replication sample's share of contributions declines faster. This is
mirrored by a two-sided rank sum test which is only significant at a
five percent level in periods three and four.

@@

How do the different contribution-paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied and redistributed).} Given that the
original sample contributed more, one would expect the respective groups
to be more wealthy. A mere mean comparison indicates just that: An
average group in the original sample accumulated about 478 tokens. In
contrast, an average group in the replication sample accumulated about
380 tokens. This difference is insignificant at conventional levels
though: A two-sided rank sum test (comparing differences between
samples) yields a p-Value of 0.1356 for the mean stock in last period of
the game.

@@

Why? Perhaps because the heterogeneity within samples and across groups
has been too large to \emph{detect} a significant difference. The right
panel of Figure~\ref{fig-growth-heterogeneity} depicts heterogeneity: In
the replication sample, the richest group earned 1425 tokens (which is
about 1781\% of the initial endowment) whereas the poorest group ends up
with 92 tokens (115\%). More broadly, the replication sample is
characterized by inequality between groups (\(SD_{Replication} =\)
336.06). The same holds true for the original sample
(\(SD_{Original} =\) 393.58). Hence, the heterogeneity across groups
does not differ between samples, which is remarkable because the
replication sample was drawn from a more heterogeneous (non-convenience
sample). Does it differ within groups?

@@

Figure~\ref{fig-kernel-generalizability} shows distributions of
contributions across both choices for both samples. The top panels
illustrate the behavior of the general population sample. The bottom
panels illustrate the behavior of the general population sample. The
left panels show the behavior in the VCA. The right panels show the
behavior in first period of the game. A visual inspection shows that (a)
mean contributions are positive in both tasks for both
samples.\footnote{The same holds true for median contributions.} (b)
Furthermore, average contributions are lower in the VCA. (c) In contrast
to \citet{GKLS2020}'s observation, we do not observe a difference
between samples in the abstract game's contribution behavior. (d)
However, the student's share of income contributed to the VCA is
significantly lower than the general population sample's contributions
(two-sided rank sum test, p=0.0085). Taken together, these aggregate
results indicate, that the consistency between tasks is higher for the
general population than it is for students. Or, to put it differently,
the general population's behavior in the abstract game better predicts
their behavior in real-world mitigation context.

>>>>

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public good game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

Today's actions are tomorrow's result. There are many settings in which
current decisions affect future outcomes and with it, future decision
spaces. Opting for environmental friendly policies today not only
reduces carbon dioxide omissions immediately but also helps us to reach
the Paris climate targets tomorrow. Deferring these policies, may not
necessarily prevent us from reaching these targets, but it makes it more
difficult in the future. Hence, today's actions (or the omission
thereof) not only affect intermediate outcomes but also the number of
paths one can choose to reach certain goals. This applies especially to
carbon dioxide emissions where its current stock will last for well over
a millennium \citep{Inman2008, CalzolariEtAl2018}.

@@

In fact, the experiment reported in this article qualifies for the label
\emph{`pre-registered'} at a first glance: Parts of the analyses were
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. In addition, I pre-registered the exact
analyses I planned to run when I designed the experiment on
\href{https://github.com/Howquez/coopUncertainty/blob/July21Replication/analysis/reports/rmd}{GitHub}.\footnote{https://github.com/Howquez/coopUncertainty/tree/July21Replication/analysis/reports/rmd
  to run the code, you need to executed the .Rmd files in this
  repository in the order that is indicated by its file names.}

At a second glance, the analysis that made the most sense after having
collected the data and that is reported here is a little different
though: Because the more representative subject pool was exhausted
earlier than expected, I recruited students which opened up a new
research direction: assessing generalizability. Hence, this article
combines both confirmatory as well as exploratory research.

Accordingly, I changed many of the analysis scripts.\footnote{The source
  code of this document contains the analysis and can be found
  \href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}{here}:
  https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}
However, because the originally planned analysis code is archived and
reproducible, these changes are transparent. Importantly, I followed a
literate programming approach \citep{Knuth_1984, AkhtarYe_2023}. Hence,
all documents needed to analyze the data and write the report stem from
the same source. This establishes consistence between the commands one
tells the statistical software to do and the explanation one tells human
beings one told the statistical software to do. As such, (a), (b) and
the transition from (a) to (b) are not only transparent but also
comprehensible.

@@

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific} reproduction\footnote{\citet{Parsons2022}
  use the term of a \emph{conceptual replication} which means the same.}
of one treatment of GMTV's dynamic public good game. The pure
reproduction re-analyzes the original data.
\protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{errors I identified} in the original paper. The scientific
reproduction, where I utilize a different sample drawn from a different
population in a different situation, is described in the following
sections.

@@

\hypertarget{sec-replication}{%
\subsection{Pre-registered GMTV Reproduction}\label{sec-replication}}

@@

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is the reproduction sample (consisting
of both students and non-students) more pro-social than the original
sample? A two-sided rank sum test reveals that it is not (p=0.3926).
Both samples contributed 10 tokens, that is, 50\% of their endowments on
average (median and mean). Moreover, both samples' initial contributions
resemble initial contributions participants usually make in the static
game with partner matching.\footnote{See Figure 3B in
  \citet[p.989]{fehrgaechter2000}, for instance.} However, in the
dynamic game presented here, we are particularly interested in the
subsequent periods because differences add up exponentially. Do the two
groups remain similar over the course of time?

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{no}.
Figure~\ref{fig-share-of-contributions} illustrates that the samples
make similar contributions at the beginning and the end of the game but
behave differently in between. More precisely, the left panel--depicting
the average contributions in absolute terms--shows that the original
sample contributed more than the reproduction sample \emph{in all but
the first and last period}.\footnote{In period five, this difference is
  significant: a two-sided rank sum test yields p=0.0486.} For this
reason, the original sample's behavior differs from the reproduction
sample's behavior in two aspects: it contributes more and exhibits a
considerable drop in the last period (whereas the reproduction sample's
contributions flatten).

Note that increasing contributions over time imply increasing endowments
over time. Hence, absolute contributions do not tell much about the
willingness to cooperate. For this reason, the right panel in
Figure~\ref{fig-share-of-contributions} shows the average \emph{share of
endowments contributed} over time. Both samples exhibit a similar
pattern: their share of endowments contributed declined and did not
stabilize. However, both samples also differ with respect to one aspect:
the reproduction sample's share of contributions declines faster. This
is mirrored by a two-sided rank sum test which is only significant at a
five percent level in periods three and four.

@@

How do the different contribution-paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied and redistributed).} Given that the
original sample contributed more, one would expect the respective groups
to be more wealthy. A mere mean comparison indicates just that: An
average group in the original sample accumulated about 478 tokens. In
contrast, an average group in the reproduction sample accumulated about
380 tokens. This difference is insignificant at conventional levels
though: A two-sided rank sum test (comparing differences between
samples) yields a p-Value of 0.1356 for the mean stock in last period of
the game.

@@

Why? Perhaps because the heterogeneity within samples and across groups
has been too large to \emph{detect} a significant difference. The right
panel of Figure~\ref{fig-growth-heterogeneity} depicts heterogeneity: In
the reproduction sample, the richest group earned 1425 tokens (which is
about 1781\% of the initial endowment) whereas the poorest group ends up
with 92 tokens (115\%). More broadly, the reproduction sample is
characterized by inequality between groups (\(SD_{Replication} =\)
336.06). The same holds true for the original sample
(\(SD_{Original} =\) 393.58). Hence, the heterogeneity across groups
does not differ between samples, which is remarkable because the
reproduction sample was drawn from a more heterogeneous (non-convenience
sample). Does it differ within groups?

@@

Figure~\ref{fig-kernel-generalizability} shows distributions of
contributions across both choices for both samples. The top panels
illustrate the behavior of the general population sample. The bottom
panels illustrate the behavior of the student sample. The left panels
show the behavior in the VCA. The right panels show the behavior in
first period of the game. A visual inspection shows that (a) mean
contributions are positive in both tasks for both samples.\footnote{The
  same holds true for median contributions.} (b) Furthermore, average
contributions are lower in the VCA. (c) In contrast to
\citet{GKLS2020}'s observation, we do not observe a difference between
samples in the abstract game's contribution behavior. (d) However, the
student's share of income contributed to the VCA is significantly lower
than the general population sample's contributions (two-sided rank sum
test, p=0.0085). Taken together, these aggregate results indicate, that
the consistency between tasks is higher for the general population than
it is for students. Or, to put it differently, the general population's
behavior in the abstract game better predicts their behavior in
real-world mitigation context.
